\section{Objection.js}
Objection.js is an ORM based on the Knex.js query builder and is closely
connected to the Knex ecosystem. It supports SQLite3, Postgres, and MySQL while
also being compatible with many other database engines supported by Knex
\cite{objectionGitHub}. Unlike Sequelize or TypeORM, Objection.js does not
provide extensive abstraction; it only offers Models that assist with query
composition, types, and relation fetching. The Models primarily serve as a
starting point for SQL queries rather than offering the complex functionality of
Active Record or Data Mapper patterns. This results in syntax which is quite
similar to Knex as shown in \autoref{lst:knexObjection}.

The package has 128,872 weekly downloads \cite{objectionNpm} and 6,972 stars on
GitHub \cite{objectionGitHub}. Objection.js predates the popularization of
TypeScript and is therefore developed fully in JavaScript. Its former lead
developer admits that the package requires a~significant rewrite of types to
compete in TypeScript support, which is not feasible due to the workload
required \cite{objectionFuture}. Objection.js was not maintained for over a
year, accumulating technical debt and leading to incompatibility with TypeScript
4.8 and newer due to enhancements to the \texttt{strictNullChecks} option. A~new
version has since been published \cite{objection302}, fixing some issues that
arose during that time. New developers have started working on maintaining the
project as of writing this thesis.

The package is released under the MIT Licence and offers high-quality web
documentation, including extensive usage guides and a~detailed API Reference
\cite{objectionDocs}. A~unique feature of Objection.js is the ability to provide
JSON Schema validation for inserted objects, ensuring database data consistency.

Objection.js has only three dependencies for full functionality: Knex, AJV, and
db-errors. The AJV dependency provides JSON Schema validation, while db-errors,
originating from the same company that published Objection.js, aims to deliver a
unified API for handling various errors produced by different database engines.
Although the db-errors package has not been updated for over four years, it is
unlikely to pose a~significant security threat to the package. Objection.js now
also supports both ESM and CommonJS, with the latest release fixing ESM
compatibility issues.

\begin{listing}
\caption{Comparison of Knex.js and Objection.js syntax for count query}
\label{lst:knexObjection}
\begin{minted}{typescript}
// Knex.js
knexInstance
  .select('stock_info')
  .from('toys_producer')
  .where('id', '=', id)
  .first()
  .then(data => data?.stock_info ?? {})

// Objection.js
ToysProducer.query()
  .select('stock_info')
  .where('id', '=', id)
  .first()
  .then(data => data?.stockInfo ?? {})
\end{minted}
\end{listing}

\subsection*{Performance in benchmarks}

Objection.js could only be included in the performance benchmarks only after its
typings were updated to be compatible with TypeScript 4.8 and later while
maintaining \texttt{strictNullChecks}, a~requirement for the proper functioning
of other packages like Zapatos. Once the compatibility issues were resolved with
manually edited types, Objection.js demonstrated performance largely on par with
other ORM packages such as MikroORM and TypeORM. However, a~notable exception
was observed in the getCatColor test, where Objection.js was significantly
slower than its counterparts.

\newpage

This performance discrepancy was traced back to the method that was used for
fetching related entities: \texttt{withGraphFetched}, the only type-supported
way to fetch relations beyond a~single relationship in Objection.js without
resorting to plain Knex query builder calls \cite{objectionDocs}. This method
was selected for type support, but it caused the speed issue encountered. The
method retrieves data through three separate, dependent database calls,
resulting in considerably longer query execution times even with minimal
latency. While this approach may offer some benefits in specific scenarios, such
as applying limits to individual queries to alleviate the load on massive
databases, it generally falls short of the performance optimization capabilities
offered by relational database management systems (RDBMS) that utilize single,
more complex queries.

Another challenge posed by Objection.js lies in its limited typing support. Many
inherited methods from Knex lack template variables to specify the result,
despite their original counterparts providing such functionality. Consequently,
developers often need to cast the results manually to ensure the correct final
type, except for the most basic queries. It was the only package which provided
completely incorrect resulting types, which forced the usage of \texttt{x as
unknown as DesiredType} casting.

The composition of Objection.js, which positions itself as not quite an ORM nor
a pure query builder, often requires developers to have direct access to the
query. However, this approach often obscures vital information, forcing
developers to rely on experience or testing to understand certain aspects of the
system. For instance, the table name is hidden when using predefined relations
for joining. Developers must determine the alias assigned to the joined table
for \texttt{WHERE} clauses without any clear guidance from the package.