\subsection{Waterline}\label{ch:waterline}

The Sails framework, an Model-View-Controller style \cite[p.
330]{fowler-patterns-2003} architecture framework for JavaScript, is designed
for building modern web applications. As part of this framework, the Model
component is implemented using an ORM called Waterline. Sails is relatively less
popular in the JavaScript ecosystem, with only 31,492 downloads, which pales in
comparison to giants like Express (28,825,585 weekly downloads
\cite{express_2022}) or Koa (1,600,532 weekly downloads \cite{koa_2023}). While
Waterline's primary purpose is to serve as a~component within the Sails
framework, it is also available as a~standalone package.

However, during the implementation process, several issues were encountered with
Waterline. The package does not provide its own types but relies on optional
types from the \texttt{@types} repository \cite{definitelytyped/types/waterline}, which
is maintained by volunteers. Unfortunately, these types are outdated, as they
only cover version 0.13.x, while the latest Waterline version is 0.15.2. As a
result, these types do not accurately reflect the package's functionality, and
initialization using them does not work. Similarly, following the documentation
for a~standalone setup, as shown on the framework's website, also proved
non-functional as the documentation \cite{waterline-docs} is written mainly for
the full framework usage, with pure waterline usage being an afterthought.

Due to these inconsistencies and the inability to overcome them, Waterline was
disqualified from the comparison.