\chapter{Benchmark database schema design}

This chapter describes the database used for performance testing of the ORM and
query builder packages. The database is designed around imaginary data
collection about cats, their home domiciles and toys found within these houses,
and the toys’ manufacturers. The database comprises six main entities – cat, cat
colours, colour hex codes, houses, toys and toy producers.

The cat entity instances represent individual cats which we want to monitor.
Each has a unique identifier, name and date of birth, all of which are nullable
except for the identifier. This entity aims to represent the basic database
table and to verify the correct handling of the data type from Postgres, as
JavaScript Date time represents a moment, including time. In contrast, the
database entry would only contain the date. Additionally, the cat entity uses
big integer data type, and handling numbers beyond the standard range allocated
in JavaScript is tested. The cat colour and colour hex code are two entities
that represent the cat colour by its name and by its hex code. The entities are
intentionally split in this way to use identifying relation – the primary key of
the hex colour entity is also a foreign key referencing the id of the cat colour
entity. 

House entity represents domiciles where the cats spend their time at their
behest. The relation must also account for ambitious cats using several houses
as their homes. The main aim is to test the difficulty of implementing and using
simple many-to-many relations. The only attribute that provides new data type or
behaviour is the simple has\_dog attribute, specified as a Boolean. It is one of
several attributes that test the frameworks’ ability to correctly type and
convert the data recovered from the database. The database types which are
tested are Date, Timestamp (with time zone), Boolean, Big Integer, JSON, JSONB
(decomposed binary representation of JSON) and Numeric (Decimal representation
of numbers). 

The houses can be equipped with many toys for the cats to use. The relation
between houses and toys is modelled through a decomposition table which contains
attributes representing the number of the same toy in the house. While the
primary keys are the identifiers of the house and toy, the decomposition with
the amount, rather than several records with an additional identifier, is
designed to test the ability to insert a record if it does not exist or update
the value referencing its previous state. If more toys are purchased, the owner
of the house does not suddenly throw out all toys they already had; they will
add them to their current pile. This operation is often called “upsert” – a
combination of update and insert, and some database engines, such as
CockroachDB, implement it explicitly under this name. PostgreSQL achieves it
using the ON CONFLICT statement in INSERT query. It also tests the handling of
composite primary keys, a standard paradigm in many databases. 

The toy entity purpose in testing is in numeric data type used in price
attribute and usage of additional column attributes such as CHECK constraints or
DEFAULT values in the column. Column naughty is focused on commonly problematic
strings in software development, such as special Unicode characters, emojis and
other issues that could come up in handling data from the database, especially
if the encoding is not correctly handled. Toys producers host the JSON columns
to test if it is possible to use advanced JSON traversal and query operators
provided in PostgreSQL (and their equivalents in other database management
systems).
